---
title: "project revised"
author: "Andrew Wrench"
output: html_document
---
```{r}
library('psych')
library('ggplot2')
library('gridExtra')
library('ggcorrplot')
library('factoextra')
library('cluster')
library('NbClust')
library('dendextend')
library('tidyr')
library('dplyr')
library('stringr')
```

```{r}
hrv_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/hrv_measurements.csv')
sleep_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/sleep.csv')
bp_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/blood_pressure.csv')
heart_rate_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/heart_rate.csv')
participants_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/participants.csv')
scales_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/scales_description.csv')
surveys_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/surveys.csv')
wearables_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/wearables.csv')
weather_data = read.csv('/Users/andrewwrench/Downloads/hrv-covid19-master/data/weather.csv')
```

# taking HRV_measurements data, and averaging all numeric columns for indidivual reportings specific to day and AM/PM
```{r}
hrv_data_sub = hrv_data
hrv_data_sub$measurement_datetime = str_sub(hrv_data$measurement_datetime, end = -7)
hrv_data_sub[c('date', 'time')] = str_split_fixed(hrv_data_sub$measurement_datetime, ' ', 2)
hrv_data_sub = subset(hrv_data_sub, select = -c(measurement_datetime, time_of_day, rr_code, how_feel, how_mood, how_sleep, tags, rr_data))
hrv_data_sub
hrv_data_sub$time = as.integer(hrv_data_sub$time)
hrv_data_sub$time = ifelse(hrv_data_sub$time >= 12, 'PM', 'AM')
hrv_data_sub
unique(hrv_data_sub$time)
hrv_data_sub_means = hrv_data_sub %>% group_by(user_code, date, time) %>% summarise(across(everything(), mean), .groups = 'drop')  %>% as.data.frame()
hrv_data_sub_means
```

# changing age ranges to single numbers, genders to t/f, and cutting out useless columns
```{r}
participants_data
unique(participants_data$age_range)
participants_data$age_range[participants_data$age_range == '18-24'] = 1
participants_data$age_range[participants_data$age_range == '25-34'] = 2
participants_data$age_range[participants_data$age_range == '35-44'] = 3
participants_data$age_range[participants_data$age_range == '45-54'] = 4
participants_data$age_range[participants_data$age_range == '55-64'] = 5
participants_data$age_range[participants_data$age_range == '65-74'] = 6
participants_data$gender[participants_data$gender == 'm'] = T
participants_data$gender[participants_data$gender == 'f'] = F
participants_data$gender = as.logical(participants_data$gender)
participants_data_sub = subset(participants_data, select = -c(city, country, symptoms_onset))
participants_data_sub
```

# joining participants and hrv datasets by user_code, date, and time so that every hrv avg value for an am/pm observation is filled with static gender, weight, etc
```{r}
hrv_participants_data = full_join(hrv_data_sub_means, participants_data_sub, by = 'user_code')
hrv_participants_data
dim(hrv_data_sub_means)
dim(hrv_participants_data)
dim(participants_data_sub)
hrv_participants_data$user_code_date_time = paste0(hrv_data_sub_means$user_code, hrv_data_sub_means$date, hrv_data_sub_means$time)
hrv_participants_data
```

# averaging bp values for each participant for am/pm per day, noticeable lose a lot of columns here
```{r}
bp_data_sub = bp_data
bp_data_sub$measurement_datetime = str_sub(bp_data_sub$measurement_datetime, end = -7)
bp_data_sub[c('date', 'time')] = str_split_fixed(bp_data_sub$measurement_datetime, ' ', 2)
bp_data_sub = subset(bp_data_sub, select = -c(measurement_datetime))
bp_data_sub$time = as.integer(bp_data_sub$time)
bp_data_sub$time = ifelse(bp_data_sub$time >= 12, 'PM', 'AM')
bp_data_sub
unique(bp_data_sub$time)
bp_data_sub_means = bp_data_sub %>% group_by(user_code, date, time) %>% summarise(across(everything(), mean), .groups = 'drop')  %>% as.data.frame()
bp_data_sub_means$user_code_date_time = paste0(bp_data_sub_means$user_code, bp_data_sub_means$date, bp_data_sub_means$time)
bp_data_sub_means
```

# joining bp table to hrv and participant tables
```{r}
bp_hrv_participants_data = full_join(hrv_participants_data, bp_data_sub_means, by = 'user_code_date_time')
bp_hrv_participant_data
dim(bp_hrv_participant_data)
length(unique(bp_hrv_participant_data$user_code_date_time))
bp_hrv_participants_data_sub = subset(bp_hrv_participants_data, select = -c(user_code.x, date.x, time.x, user_code.y, date.y, time.y))
bp_hrv_participants_data_sub
```

#changing wearables data to better format
```{r}
wearables_data_sub = wearables_data
wearables_data_sub_means = wearables_data_sub %>% group_by(user_code, day) %>% summarise(across(everything(), mean), .groups = 'drop')  %>% as.data.frame()
wearables_data_sub_means$user_code_date_time = paste0(wearables_data_sub_means$user_code, wearables_data_sub_means$day, 'AM')
wearables_data_sub_means
```

# joining wearables to bp, hrv, participants table
```{r}
wearables_bp_hrv_participants_data = full_join(bp_hrv_participants_data, wearables_data_sub_means, by = 'user_code_date_time')
wearables_bp_hrv_participants_data
dim(wearables_bp_hrv_participants_data)
length(unique(wearables_bp_hrv_participants_data$user_code_date_time))
wearables_bp_hrv_participants_data_sub = subset(wearables_bp_hrv_participants_data, select = -c(user_code.x, date.x, time.x, user_code.y, date.y, time.y))
wearables_bp_hrv_participants_data_sub
dim(bp_hrv_participant_data)
```

# similar conversion on heart rate data for new avg
```{r}
heart_rate_data_sub = heart_rate_data
heart_rate_data_sub$datetime = str_sub(heart_rate_data_sub$datetime, end = -7)
heart_rate_data_sub[c('date', 'time')] = str_split_fixed(heart_rate_data_sub$datetime, ' ', 2)
heart_rate_data_sub = subset(heart_rate_data_sub, select = -c(is_resting, datetime))
heart_rate_data_sub$time = as.integer(heart_rate_data_sub$time)
heart_rate_data_sub$time = ifelse(heart_rate_data_sub$time >= 12, 'PM', 'AM')
heart_rate_data_sub
unique(heart_rate_data_sub$time)
heart_rate_data_sub_means = heart_rate_data_sub %>% group_by(user_code, date, time) %>% summarise(across(everything(), mean), .groups = 'drop')  %>% as.data.frame()
heart_rate_data_sub_means$user_code_date_time = paste0(heart_rate_data_sub_means$user_code, heart_rate_data_sub_means$date, heart_rate_data_sub_means$time)
heart_rate_data_sub_means
```

# full join of heart rate data with wearables, bp, hrv, participant data
```{r}
heart_rate_wearables_bp_hrv_participants_data = full_join(wearables_bp_hrv_participants_data_sub, heart_rate_data_sub_means, by = 'user_code_date_time')
heart_rate_wearables_bp_hrv_participants_data
dim(heart_rate_wearables_bp_hrv_participants_data)
length(unique(heart_rate_wearables_bp_hrv_participants_data$user_code_date_time))
heart_rate_wearables_bp_hrv_participants_data_sub = subset(heart_rate_wearables_bp_hrv_participants_data, select = -c(user_code.x, user_code.y, day))
heart_rate_wearables_bp_hrv_participants_data_sub
dim(wearables_bp_hrv_participants_data)
```

# sleep data conversion
```{r}
sleep_data_sub = sleep_data
sleep_data_sub$sleep_begin = str_sub(sleep_data_sub$sleep_begin, end = -7)
sleep_data_sub
sleep_data
summary(sleep_data)
sleep_data_sub[c('date', 'time')] = str_split_fixed(sleep_data_sub$sleep_begin, ' ', 2)
sleep_data_sub = subset(sleep_data_sub, select = -c(day, sleep_end))
sleep_data_sub$time = as.integer(sleep_data_sub$time)
sleep_data_sub$time = ifelse(sleep_data_sub$time >= 12, 'PM', 'AM')
sleep_data_sub
unique(heart_rate_data_sub$time)
sleep_data_sub_means = sleep_data_sub %>% group_by(user_code, date, time) %>% summarise(across(everything(), mean), .groups = 'drop')  %>% as.data.frame()
sleep_data_sub_means$user_code_date_time = paste0(sleep_data_sub_means$user_code, sleep_data_sub_means$date, sleep_data_sub_means$time)
sleep_data_sub_means
```

# adding sleep data to heart rate, wearables, bp, hrv, and participants data
```{r}
sleep_heart_rate_wearables_bp_hrv_participants_data = full_join(heart_rate_wearables_bp_hrv_participants_data_sub, sleep_data_sub_means, by = 'user_code_date_time')
sleep_heart_rate_wearables_bp_hrv_participants_data
dim(sleep_heart_rate_wearables_bp_hrv_participants_data)
sleep_heart_rate_wearables_bp_hrv_participants_data_sub = subset(sleep_heart_rate_wearables_bp_hrv_participants_data, select = -c(time.x, date.x, date.y, time.y, pulse_min.y, pulse_max.y, pulse_average.y, pulse_min.x, pulse_max.x, pulse_average.x))
sleep_heart_rate_wearables_bp_hrv_participants_data_sub
```

# organizing surveys data
```{r}
surveys_data_sub = surveys_data
surveys_data_sub = subset(surveys_data_sub, select = -c(text))
surveys_data_sub
surveys_data_sub_median = surveys_data_sub %>% group_by(user_code, scale, created_at) %>% summarise(across(everything(), median), .groups = 'drop')  %>% as.data.frame()
surveys_data_sub_median$user_code_date_time = paste0(surveys_data_sub_median$user_code, surveys_data_sub_median$created_at, 'AM')
surveys_data_sub_median
```

# joining surveys with other data tables
```{r}
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data = full_join(sleep_heart_rate_wearables_bp_hrv_participants_data_sub, surveys_data_sub_median, by = 'user_code_date_time')
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data
dim(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data)
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub = subset(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data, select = -c(user_code.y, user_code.x))
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub
```

```{r}
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric = select_if(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub, is.numeric)
summary(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric)
dim(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric)
```

```{r}
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric_scaled = scale(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric)
summary(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric)
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric_scaled
surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric_scaled_pca = prcomp(surveys_sleep_heart_rate_wearables_bp_hrv_participants_data_sub_numeric_scaled)
```