---
title: "ordinal_regression"
output: html_document
date: '2023-04-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Resources

https://stats.oarc.ucla.edu/r/dae/ordinal-logistic-regression/

https://stats.stackexchange.com/questions/493254/why-ordinal-target-in-classification-problems-need-special-attention

# Libraries
```{r}
library(tidyverse)
library(MASS)
library(ROCR)
library(ordinalForest)
library(psych)
```

# Read data
```{r}
surveys_wide_join_train_data_clean <- read_tsv("../data_clean/surveys_wide_join_train_data_clean.tsv")
```

```{r}
surveys_wide_join_test_data_clean <- read_tsv("../data_clean/surveys_wide_join_test_data_clean.tsv")
```

Remove observation ids
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(-outcome_user_code, -outcome_created_at)
```

Re-code  outcome as ordered 
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate(outcome_S_COVID_OVERALL = as.ordered(outcome_S_COVID_OVERALL))
```

Drop binary outcome version
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(-outcome_symptom_severity)

colnames(surveys_wide_join_train_data_clean)
```

Recode how feel answers as ordinal
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate_at(vars(contains("how")), round) %>%
  mutate_at(vars(contains("how")), as.ordered)
```

Scale numeric variables
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate_if(is.numeric, scale)
```

```{r}
surveys_wide_join_train_data_clean$participants_gender <- as.factor(surveys_wide_join_train_data_clean$participants_gender)

#synthetic_data <- SMOTE(X = surveys_wide_join_train_data_clean %>%
        #select_if(is.numeric), target = #surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL,dup_size = 1)
#synthetic_data$syn_data$class
```


# Basic data exploration

Distributions of numeric variables

```{r}
surveys_wide_join_train_data_clean %>%
  select_if(is.numeric) %>%
  pivot_longer(cols = everything(), names_to = "feature", values_to = "value") %>%
  ggplot(aes(value)) + 
  geom_histogram() +
  facet_wrap(~feature,scales = "free")
```

```{r}
surveys_wide_join_train_data_clean %>%
  select_if(is.numeric) %>%
  psych::pairs.panels()
```

Class imbalance
```{r}
surveys_wide_join_train_data_clean %>%
  ggplot(aes(outcome_S_COVID_OVERALL)) +
  geom_histogram(stat = "count", fill = "darkblue") +
  theme_classic() +
  theme(aspect.ratio = 1) +
  ggtitle("Counts per class (ordinal)")
```




# Ordered logistic regression

Conducting an ordered logistic regression with the polyr function from the MASS package

```{r}
m_log_regression1 <- polr(outcome_S_COVID_OVERALL ~ ., data = surveys_wide_join_train_data_clean, Hess=TRUE)
```


# Predict classes

```{r}
predictions_train_m_log_regression1 <- predict(m_log_regression1,newdata = surveys_wide_join_train_data_clean)
```

MSE
https://stats.stackexchange.com/questions/338904/measures-of-ordinal-classification-error-for-ordinal-regression

Note that classes are imbalanced, so mean squared error might not work as well

```{r}
mse <- function(predictions,labels) {
  mean((as.numeric(predictions) - as.numeric(labels))^2)
}

mse_predictions_train_m_log_regression1 <- mse(predictions_train_m_log_regression1, surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)
```


# LOOCV 

Calculate mse from LOOCV

```{r}
# Function to fit model and predict train labels, returns mse 
get_loocv_polr_mse <- function(f, d, l) {
  m <- polr(f, d, Hess=TRUE)
  m_predictions <- predict(m,newdata = d)
  
  return(mse(predictions = m_predictions, labels = l))
} 

# For loop where we skip one observation at a time, returns list with mse
mse_loocv <- as.list(1:nrow(surveys_wide_join_train_data_clean))
for (i in 1:nrow(surveys_wide_join_train_data_clean)) {
    mse_loocv[[i]] <- try(get_loocv_polr_mse(f = outcome_S_COVID_OVERALL ~ ., d = surveys_wide_join_train_data_clean[-i,], l = surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL[-i] ))
   
}

# Take the mean mse overl all LOO trials
mse_loocv <- mse_loocv[map_lgl(mse_loocv, ~ class(.x) != "try-error")]
mse_loocv_est <- mean(unlist(mse_loocv))

```


Calculate J from LOOCV

```{r}
# Function to fit model and predict train labels, returns mse 
get_loocv_polr_j <- function(f, d, l) {
  m <- polr(f, d, Hess=TRUE)
  m_predictions <- predict(m,newdata = d)
  
  return(perff_equal(ytest = as.factor(as.character(l)),ytestpred = m_predictions))
} 

# For loop where we skip one observation at a time, returns list with mse
j_loocv <- as.list(1:nrow(surveys_wide_join_train_data_clean))
for (i in 1:nrow(surveys_wide_join_train_data_clean)) {
    j_loocv[[i]] <- try(get_loocv_polr_j(f = outcome_S_COVID_OVERALL ~ ., d = surveys_wide_join_train_data_clean[-i,], l = surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL[-i] ))
   
}

# Take the mean mse overl all LOO trials
j_loocv <- j_loocv[map_lgl(j_loocv, ~ class(.x) != "try-error")]
j_loocv_est <- mean(unlist(j_loocv))
j_loocv_est

summary(unlist(j_loocv))
```

# AIC

```{r}
stepAIC(m_log_regression1)

exp(coef(m_log_regression1))
```


# Subset selection

```{r}
m_log_regression1_step <- stepAIC(m_log_regression1,direction = "both",score = list(lower = ~1,upper=~.))
```

Compare full and subset model 
```{r}
summary(m_log_regression1_step)
summary(m_log_regression1)

anova(m_log_regression1, m_log_regression1_step)
```

Although the models are not significantly different according to the LRT, the step model has lower AIC.

Variables:
```{r}
sort(abs(m_log_regression1_step$coefficients[-c(10:13)]),decreasing = TRUE) %>%
  enframe() %>%
  mutate(name = fct_reorder(as.factor(name),.x = value)) %>%
  ggplot(aes(value, name, fill = value)) +
  geom_col() +
  ylab("") +
  theme_classic() +
  theme(aspect.ratio = 1, legend.position = "none") +
  xlab("abs(coefficient)") +
  ggtitle("Ordered logistic regression coefficients \n
          after step selection")

```





# Try Ordinal Forest

```{r}


surveys_wide_join_train_data_clean <- as.data.frame(surveys_wide_join_train_data_clean)



m_ordfor_1 <- ordfor(depvar = "outcome_S_COVID_OVERALL", surveys_wide_join_train_data_clean %>%
  mutate_if(is.ordered, as.factor),perffunction = "probability")

enframe(m_ordfor_1$varimp) %>%
  mutate(name = fct_reorder(as.factor(name),.x = value)) %>%
  ggplot(aes(value, name, fill = value)) +
  geom_col() +
  ylab("") +
  theme_classic() +
  theme(aspect.ratio = 1, legend.position = "none") +
  xlab("Ranked Probability Score") +
  ggtitle("Ordinal Forest Variable Importance")

```


Explore important variables:

```{r}
surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_resting_pulse,participants_weight, color = outcome_S_COVID_OVERALL)) +
  geom_point() +
  theme_linedraw() +
  theme(aspect.ratio = 1)

surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_pulse_min,hrv_vlf, color = outcome_S_COVID_OVERALL)) +
  geom_point() +
  theme_linedraw() +
  theme(aspect.ratio = 1, legend.position = "bottom") 
```


```{r}
surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_resting_pulse,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_total_calories_burned,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(hrv_vlf,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(weather_humidity,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(participants_age_range,outcome_S_COVID_OVERALL)) +
  geom_bin_2d()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(participants_gender,outcome_S_COVID_OVERALL)) +
  geom_bin_2d()
```
```{r}
surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_resting_pulse,outcome_S_COVID_OVERALL)) +
  geom_boxplot(aes(color = outcome_S_COVID_OVERALL)) +
  theme_classic() +
  theme(aspect.ratio = 1)

surveys_wide_join_train_data_clean %>%
  ggplot(aes(hrv_vlf,outcome_S_COVID_OVERALL)) +
  geom_boxplot(aes(color = outcome_S_COVID_OVERALL)) +
  theme_classic() +
  theme(aspect.ratio = 1)


```


"Youden’s J statistic is calculated with respect to each
class ("observation/prediction in class j" vs. "observation/prediction NOT in class j" (j=1,...,J)) and
the simple average of the J results taken."

```{r}

predictions_m_ordfor_1 <- predict(m_ordfor_1,newdata = surveys_wide_join_train_data_clean)
predictions_m_log_regression1_step <- predict(m_log_regression1_step,newdata = surveys_wide_join_train_data_clean)
predictions_m_log_regression1 <- predict(m_log_regression1,newdata = surveys_wide_join_train_data_clean)

perff_equal(ytest = as.factor(as.character(surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_ordfor_1$ypred)

perff_equal(ytest = as.factor(as.character(surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_log_regression1_step)

perff_equal(ytest = as.factor(as.character(surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_log_regression1)


```


# Add test data

Process test data

```{r}
surveys_wide_join_test_data_clean <- read_tsv("../data_clean/surveys_wide_join_test_data_clean.tsv")
```

Remove observation ids
```{r}
surveys_wide_join_test_data_clean <- surveys_wide_join_test_data_clean %>%
  dplyr::select(-outcome_user_code, -outcome_created_at)
```

Re-code  outcome as ordered 
```{r}
surveys_wide_join_test_data_clean <- surveys_wide_join_test_data_clean %>%
  mutate(outcome_S_COVID_OVERALL = as.ordered(outcome_S_COVID_OVERALL))
```

Drop binary outcome version
```{r}
surveys_wide_join_test_data_clean <- surveys_wide_join_test_data_clean %>%
  dplyr::select(-outcome_symptom_severity)

colnames(surveys_wide_join_test_data_clean)
```

Recode how feel answers as ordinal
```{r}
surveys_wide_join_test_data_clean <- surveys_wide_join_test_data_clean %>%
  mutate_at(vars(contains("how")), round) %>%
  mutate_at(vars(contains("how")), as.ordered)
```

Scale numeric variables
```{r}
surveys_wide_join_test_data_clean <- surveys_wide_join_test_data_clean %>%
  mutate_if(is.numeric, scale)
```

# Test models


```{r}
predictions_m_ordfor_1 <- predict(m_ordfor_1,newdata = surveys_wide_join_test_data_clean)
predictions_m_log_regression1_step <- predict(m_log_regression1_step,newdata = surveys_wide_join_test_data_clean)
predictions_m_log_regression1 <- predict(m_log_regression1,newdata = surveys_wide_join_test_data_clean)




```


## Explore J 

```{r}

perff_test <- data.frame(model = c("ordinalForest","logistic_step","logistic_full"),
  j = c(perff_equal(ytest = as.factor(as.character(surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_ordfor_1$ypred),
          perff_equal(ytest = as.factor(as.character(surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_log_regression1_step),
          perff_equal(ytest = as.factor(as.character(surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_log_regression1)))

perff_test %>%
  mutate(model = fct_reorder(as.factor(model),j)) %>%
  ggplot(aes(j, model)) +
  geom_col(aes(fill = j)) +
  theme_classic() +
  theme(aspect.ratio = 1, legend.position = "none") +
  ylab(NULL) +
  xlab("J") +
  ggtitle("Test J comparison")

```

# Calculate test RMSE

```{r}


mse_test <- data.frame(model = c("ordinalForest","logistic_step","logistic_full"),
  mse = c(mse(predictions = predictions_m_ordfor_1$ypred, labels = surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL),
                   mse(predictions = predictions_m_log_regression1_step, labels = surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL),
                   mse(predictions = predictions_m_log_regression1, labels = surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL)))

mse_test %>%
  mutate(mse = sqrt(mse)) %>%
  ggplot(aes(mse, model)) +
  geom_col(aes(fill = mse)) +
  theme_classic() +
  theme(aspect.ratio = 1, legend.position = "none") +
  ylab(NULL) +
  xlab("RMSE") +
  ggtitle("Test RMSE comparison") +
  geom_text(nudge_x = 0.2, aes(label = round(mse,2)))

```




Convert predictions to binary

```{r}

n_cutoffs <- length(levels(predictions_m_ordfor_1$ypred))

predictions_m_ordfor_1_binary_list <- map(levels(predictions_m_ordfor_1$ypred),
    ~ as.numeric(as.character(predictions_m_ordfor_1$ypred) == .x))

predictions_m_log_regression1_step_list <- map(levels(predictions_m_log_regression1_step),
    ~ as.numeric(as.character(predictions_m_log_regression1_step) == .x))

predictions_m_log_regression1_list <- map(levels(predictions_m_log_regression1),
    ~ as.numeric(as.character(predictions_m_log_regression1) == .x))

outcome_S_COVID_OVERALL_list <- map(levels(surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL),
    ~ as.numeric(as.character(surveys_wide_join_test_data_clean$outcome_S_COVID_OVERALL) == .x))



```


Calculate ROC AUC

```{r}
predictions_m_ordfor_1_binary_list_auc <- map2_dbl(predictions_m_ordfor_1_binary_list,outcome_S_COVID_OVERALL_list,
     function(.x,.y){
       performance(prediction(predictions = as.numeric(.x),labels = as.numeric(.y)),
           measure = "auc")@y.values[[1]]
     })

predictions_m_log_regression1_step_list_auc <- map2_dbl(predictions_m_log_regression1_step_list,outcome_S_COVID_OVERALL_list,
     function(.x,.y){
       performance(prediction(predictions = as.numeric(.x),labels = as.numeric(.y)),
           measure = "auc")@y.values[[1]]
     })

predictions_m_log_regression1_list_auc <- map2_dbl(predictions_m_log_regression1_list,outcome_S_COVID_OVERALL_list,
     function(.x,.y){
       performance(prediction(predictions = as.numeric(.x),labels = as.numeric(.y)),
           measure = "auc")@y.values[[1]]
     })

outcome_S_COVID_OVERALL_list_auc <- map2_dbl(outcome_S_COVID_OVERALL_list,outcome_S_COVID_OVERALL_list,
     function(.x,.y){
       performance(prediction(predictions = as.numeric(.x),labels = as.numeric(.y)),
           measure = "auc")@y.values[[1]]
     })

```

# Calculate per class AUC

```{r}
per_class_auc <- data.frame("auc" = c(predictions_m_ordfor_1_binary_list_auc,
                     predictions_m_log_regression1_step_list_auc,
                     predictions_m_log_regression1_list_auc),
           "class"= as.factor(rep(1:5, 3)),
           "model" = rep(c("ordinalForest","logistic_step","logistic_full"),each = 5)) 
mean_auc <- per_class_auc %>%
  group_by(model) %>%
  summarise(auc = mean(auc))

per_class_auc %>%
  ggplot(aes(auc,model, color = class)) +
  geom_point(size = 3) +
  geom_point(data = mean_auc, size = 5, color = "red", shape = 4) +
  theme_light() +
  scale_color_viridis_d() +
  ylab(NULL) +
  xlab("AUC") +
  theme(aspect.ratio = 1) +
  ggtitle("Per class AUC")


```





