---
title: "ordinal_regression"
output: html_document
date: '2023-04-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Aims

# Resources

https://stats.oarc.ucla.edu/r/dae/ordinal-logistic-regression/

https://stats.stackexchange.com/questions/493254/why-ordinal-target-in-classification-problems-need-special-attention

# Libraries
```{r}
library(tidyverse)
library(MASS)
library(ROCR)
```

# Read data
```{r}
surveys_wide_join_train_data_clean <- read_tsv("../data_clean/surveys_wide_join_train_data_clean.tsv")
```


Remove observation ids
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(-outcome_user_code, -outcome_created_at)
```

Re-code  outcome as ordered 
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate(outcome_S_COVID_OVERALL = as.ordered(outcome_S_COVID_OVERALL))
```

Drop binary outcome version
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(-outcome_symptom_severity)

colnames(surveys_wide_join_train_data_clean)
```

Recode how feel answers as ordinal
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate_at(vars(contains("how")), round) %>%
  mutate_at(vars(contains("how")), as.ordered)
```

Scale numeric variables
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate_if(is.numeric, scale)
```

# Basic data exploration

Distributions of numeric variables

```{r}
surveys_wide_join_train_data_clean %>%
  select_if(is.numeric) %>%
  pivot_longer(cols = everything(), names_to = "feature", values_to = "value") %>%
  ggplot(aes(value)) + 
  geom_histogram() +
  facet_wrap(~feature,scales = "free")
```



# Ordered logistic regression

Conducting an ordered logistic regression with the polyr function from the MASS package

```{r}
m_log_regression1 <- polr(outcome_S_COVID_OVERALL ~ ., data = surveys_wide_join_train_data_clean, Hess=TRUE)
```


# Predict classes

```{r}
predictions_train_m_log_regression1 <- predict(m_log_regression1,newdata = surveys_wide_join_train_data_clean)
```

MSE
https://stats.stackexchange.com/questions/338904/measures-of-ordinal-classification-error-for-ordinal-regression

Note that classes are imbalanced, so mean squared error might not work as well

```{r}
mse <- function(predictions,labels) {
  mean((as.numeric(predictions) - as.numeric(labels))^2)
}

mse_predictions_train_m_log_regression1 <- mse(predictions_train_m_log_regression1, surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)
```


# LOOCV 

Calculate mse from LOOCV

```{r}
# Function to fit model and predict train labels, returns mse 
get_loocv_polr_mse <- function(f, d, l) {
  m <- polr(f, d, Hess=TRUE)
  m_predictions <- predict(m,newdata = d)
  
  return(mse(predictions = m_predictions, labels = l))
} 

# For loop where we skip one observation at a time, returns list with mse
mse_loocv <- as.list(1:nrow(surveys_wide_join_train_data_clean))
for (i in 1:nrow(surveys_wide_join_train_data_clean)) {
    mse_loocv[[i]] <- try(get_loocv_polr_mse(f = outcome_S_COVID_OVERALL ~ ., d = surveys_wide_join_train_data_clean[-i,], l = surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL[-i] ))
   
}

# Take the mean mse overl all LOO trials
mse_loocv <- mse_loocv[map_lgl(mse_loocv, ~ class(.x) != "try-error")]
mse_loocv_est <- mean(unlist(mse_loocv))

```


Calculate J from LOOCV

```{r}
# Function to fit model and predict train labels, returns mse 
get_loocv_polr_j <- function(f, d, l) {
  m <- polr(f, d, Hess=TRUE)
  m_predictions <- predict(m,newdata = d)
  
  return(perff_equal(ytest = as.factor(as.character(l)),ytestpred = m_predictions))
} 

# For loop where we skip one observation at a time, returns list with mse
j_loocv <- as.list(1:nrow(surveys_wide_join_train_data_clean))
for (i in 1:nrow(surveys_wide_join_train_data_clean)) {
    j_loocv[[i]] <- try(get_loocv_polr_j(f = outcome_S_COVID_OVERALL ~ ., d = surveys_wide_join_train_data_clean[-i,], l = surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL[-i] ))
   
}

# Take the mean mse overl all LOO trials
j_loocv <- j_loocv[map_lgl(j_loocv, ~ class(.x) != "try-error")]
j_loocv_est <- mean(unlist(j_loocv))
j_loocv_est

summary(unlist(j_loocv))
```

# AIC

```{r}
stepAIC(m_log_regression1)

exp(coef(m_log_regression1))
```


# Subset selection

```{r}
m_log_regression1_step <- stepAIC(m_log_regression1,direction = "both",score = list(lower = ~1,upper=~.))
```

Compare full and subset model 
```{r}
summary(m_log_regression1_step)
summary(m_log_regression1)

anova(m_log_regression1, m_log_regression1_step)
```

Although the models are not significantly different according to the LRT, the step model has lower AIC.

```{r}
brant::brant(m_log_regression1_step)
```




# Try Ordinal Forest

```{r}
library(ordinalForest)

data(hearth)

surveys_wide_join_train_data_clean <- as.data.frame(surveys_wide_join_train_data_clean)



m_ordfor_1 <- ordfor(depvar = "outcome_S_COVID_OVERALL", surveys_wide_join_train_data_clean %>%
  mutate_if(is.ordered, as.factor))

enframe(m_ordfor_1$varimp) %>%
  mutate(name = fct_reorder(as.factor(name),.x = value)) %>%
  ggplot(aes(value, name, fill = value)) +
  geom_col() +
  ylab("")
```


```{r}
surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_pulse_min,wearables_total_calories_burned, color = outcome_S_COVID_OVERALL)) +
  geom_point() 
```


```{r}
surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_pulse_min,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(wearables_total_calories_burned,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(hrv_vlf,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(weather_humidity,outcome_S_COVID_OVERALL)) +
  geom_boxplot()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(participants_age_range,outcome_S_COVID_OVERALL)) +
  geom_bin_2d()

surveys_wide_join_train_data_clean %>%
  ggplot(aes(participants_gender,outcome_S_COVID_OVERALL)) +
  geom_bin_2d()
```


"Youden’s J statistic is calculated with respect to each
class ("observation/prediction in class j" vs. "observation/prediction NOT in class j" (j=1,...,J)) and
the simple average of the J results taken."

```{r}

predictions_m_ordfor_1 <- predict(m_ordfor_1,newdata = surveys_wide_join_train_data_clean)
predictions_m_log_regression1_step <- predict(m_log_regression1_step,newdata = surveys_wide_join_train_data_clean)
predictions_m_log_regression1 <- predict(m_log_regression1,newdata = surveys_wide_join_train_data_clean)

perff_equal(ytest = as.factor(as.character(surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_ordfor_1$ypred)

perff_equal(ytest = as.factor(as.character(surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_log_regression1_step)

perff_equal(ytest = as.factor(as.character(surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)),ytestpred = predictions_m_log_regression1)


```







