---
title: "ordinal_regression"
output: html_document
date: '2023-04-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(tidyverse)
library(MASS)
```

# Read data
```{r}
surveys_wide_join_train_data <- read_tsv("../data_clean/surveys_wide_join_train_data.tsv")
```

# Pre-process

Remove variables not to use in the model 
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data %>%
  dplyr::select(-outcome_symptom_severity, - participants_symptoms_onset, - participants_country, - participants_city)
```

Recode outcome to ordered
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate(outcome_S_COVID_OVERALL = as.ordered(outcome_S_COVID_OVERALL))
```


Explore NA values
```{r}
na_proportion_per_feature <- (map_dbl(surveys_wide_join_train_data_clean, function(x) sum(is.na(x))) / nrow(surveys_wide_join_train_data_clean)) 


na_proportion_per_feature %>%
  enframe() %>%
  ggplot(aes(value)) +
  geom_histogram() +
  theme_light() +
  theme(aspect.ratio = 1) +
  ggtitle("Proportion of NA values per feature")
```

Remove features with > 60% NA values
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  select_if(na_proportion_per_feature < 0.6)
```

Just use mean imputation for now

Start with per person mean imputation

```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  group_by(outcome_user_code) %>%
  mutate_if(is.numeric,~ ifelse(is.na(.),mean(., na.rm = TRUE),.)) 
```

If a person does not have a value, then impute by global mean
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  ungroup() %>%
  mutate_if(is.numeric,~ ifelse(is.nan(.),mean(., na.rm = TRUE),.)) 
```

If a person does not have a value, then impute by global mean
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  mutate_if(is.numeric,~ ifelse(is.infinite(.),mean(., na.rm = TRUE),.)) 

```

Many inf values in rr data, so drop for now
```{r}
sapply(surveys_wide_join_train_data_clean,function(x) any(is.infinite(x)))

```

```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(-hrv_rr_data)

```

Correlations between variables
```{r}
corrplot::corrplot(cor(surveys_wide_join_train_data_clean %>%
      select_if(is.numeric)))
```

REmove correlated vars 

```{r}
hrv_vars_to_keep <- surveys_wide_join_train_data_clean %>%
  dplyr::select(hrv_total_power,hrv_how_feel,hrv_how_mood, hrv_how_sleep, hrv_vlf, hrv_lfhf)

surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(!starts_with("hrv"))
surveys_wide_join_train_data_clean <- cbind(surveys_wide_join_train_data_clean, hrv_vars_to_keep)

```


REmove observation ids
```{r}
surveys_wide_join_train_data_clean <- surveys_wide_join_train_data_clean %>%
  dplyr::select(-outcome_user_code, -outcome_created_at)
```


# Ordered logistic regression

Conducting an ordered logistic regression with the polyr function from the MASS package

```{r}
m <- polr(outcome_S_COVID_OVERALL ~ ., data = surveys_wide_join_train_data_clean, Hess=TRUE)

```

# Predict classes

```{r}
table(predict(m,newdata = surveys_wide_join_train_data_clean), surveys_wide_join_train_data_clean$outcome_S_COVID_OVERALL)
```

