---
title: "join_tables"
output: html_document
date: '2023-03-28'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Read data

```{r cars}
# Join tables
library(tidyverse)
data_path <- "../../hrv-covid19/data/"
file_names <- list.files(data_path)

hrv_covid19_tables <- map(paste0(data_path, file_names), read_csv)

names(hrv_covid19_tables) <- file_names

scales_description <- hrv_covid19_tables$scales_description.csv

# remove scales descriptions from main tables
hrv_covid19_tables$scales_description.csv <- NULL

```

# Clean data before joins

```{r}
# Some observations are duplicated
hrv_covid19_tables_surveys_duplicates <- hrv_covid19_tables$surveys.csv %>%
    dplyr::group_by(user_code, created_at, scale) %>%
    dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
    dplyr::filter(n > 1L)


surveys_clean <- hrv_covid19_tables$surveys.csv %>%
  filter(!user_code %in% hrv_covid19_tables_surveys_duplicates$user_code)

# Pivot wider for surveys table
surveys_wide <- surveys_clean %>%
  pivot_wider(id_cols = c("user_code","created_at"),names_from = "scale",values_from = "value",)

# Split date into separate day and time variables, and recode day variables as dates
hrv_measurements_clean <- hrv_covid19_tables$hrv_measurements.csv %>%
  separate(measurement_datetime, into = c("day","time"), sep= " ") %>%
  mutate(day = as.Date(day))


heart_rate_clean <- hrv_covid19_tables$heart_rate.csv %>%
  separate(datetime, into = c("day","time"), sep= " ") %>%
  mutate(day = as.Date(day))


blood_pressure_clean <- hrv_covid19_tables$blood_pressure.csv %>%
  separate(measurement_datetime, into = c("day","time"), sep= "  ") %>%
  mutate(day = as.Date(day))

```

# Summarize data before joining

## Pre-process outcome variable 

Select only S_COVID_OVERALL as outcome variable
```{r}
surveys_wide_outcome <- surveys_wide %>%
  select(user_code, created_at, S_COVID_OVERALL)
```

Filter out observations without user data, and users without symptoms
```{r}
surveys_wide_symptomatic_outcome <- surveys_wide_outcome %>%
  drop_na() %>%
  filter(S_COVID_OVERALL > 1)
```

Re-code outcome variable from ordinal scale, to binary severe / not severe classes

```{r}
surveys_wide_symptomatic_outcome <- 
  surveys_wide_symptomatic_outcome %>%
  mutate(symptom_severity = ifelse(S_COVID_OVERALL >= 5, "severe","not_severe"))

surveys_wide_symptomatic_outcome
```

## Pre-process hrv_measurements table 

Drop columns we won't use
```{r}
hrv_measurements_clean <- hrv_measurements_clean %>%
  select(-rr_code,-time,-time_of_day, -tags)
```

Summarize hrv_measurements at the user / day level

```{r}
hrv_measurements_clean_summarised <- hrv_measurements_clean %>%
  group_by(user_code, day) %>%
  summarise_all(mean, na.rm=TRUE)

hrv_measurements_clean_summarised$how_sleep[is.nan(hrv_measurements_clean_summarised$how_sleep)] <- NA
```

## Pre-process heart_rate table 
Drop columns we won't use
```{r}
heart_rate_clean <- heart_rate_clean %>%
  select(-time)
```

Summarize hrv_measurements at the user / day level

```{r}
heart_rate_clean_summarised <- heart_rate_clean %>%
  group_by(user_code, day, is_resting) %>%
  summarise(heart_rate = mean(heart_rate)) %>%
  pivot_wider(id_cols = c("user_code","day"),
              values_from = "heart_rate",
              names_from = "is_resting",
              names_prefix = "heart_rate_is_resting_")

```

## Pre-process blood_pressure table 

Drop unused variables

```{r}
blood_pressure_clean <- blood_pressure_clean %>%
  select(-time)
```

Summarize blood_pressure at the user / day level

```{r}
blood_pressure_clean_summarised <- blood_pressure_clean %>%
  group_by(user_code, day) %>%
  summarise_all(mean, na.rm=TRUE) %>%
  mutate_if(is.numeric, function(x) ifelse(is.nan(x),NA,x))

```

## Pre-process weather

This table is already summarized at the user / day level. Just double check that we have unique observations for each user / day pair.

```{r}
weather_nrow <- nrow(hrv_covid19_tables$weather.csv)
unique_nrow <- hrv_covid19_tables$weather.csv %>%
  select(user_code,day) %>%
  unique() %>%
  nrow()

weather_nrow == unique_nrow
```

## Pre-process wearables

This table is already summarized at the user / day level. Just double check that we have unique observations for each user / day pair.

```{r}
wearables_nrow <- nrow(hrv_covid19_tables$wearables.csv)
wearables_unique_nrow <- hrv_covid19_tables$wearables.csv %>%
  select(user_code,day) %>%
  unique() %>%
  nrow()

wearables_nrow == wearables_unique_nrow
```

## Pre-process sleep
```{r}
hrv_covid19_tables$sleep.csv
```
```{r}
sleep_nrow <- nrow(hrv_covid19_tables$sleep.csv)
sleep_unique_nrow <- hrv_covid19_tables$sleep.csv %>%
  select(user_code,day) %>%
  unique() %>%
  nrow()

sleep_nrow == sleep_unique_nrow
```

Which are duplicated?
```{r}
unique(hrv_covid19_tables$sleep.csv)

sleep_duplicated <- hrv_covid19_tables$sleep.csv %>%
  select(user_code, day) %>%
  duplicated()

sleep_dup_users_day <- hrv_covid19_tables$sleep.csv %>%
  filter(sleep_duplicated) %>%
  select(user_code, day) %>%
  mutate(user_code_day = paste(user_code, day))

hrv_covid19_tables$sleep.csv %>%
  mutate(user_code_day = paste(user_code, day))  %>%
  filter(user_code_day %in% sleep_dup_users_day$user_code_day)
  
```
Some people have different sleep_begin and sleep_end values, but since we are not using this info, just remove the columns and get unique rows. 

```{r}
sleep_clean <- hrv_covid19_tables$sleep.csv %>%
  select(-sleep_begin, -sleep_end) %>%
  unique()
```

Double check that user code / day are unique

```{r}
sleep_nrow <- nrow(sleep_clean)
sleep_unique_nrow <- sleep_clean %>%
  select(user_code,day) %>%
  unique() %>%
  nrow()

sleep_nrow == sleep_unique_nrow
```

## Pre-process participants

Participants are already summarized at the user/day level. Just double check that observations are unique.

```{r}
nrow(hrv_covid19_tables$participants.csv) == nrow(unique(hrv_covid19_tables$participants.csv))
```

# Join tables

Before joining, change data columns to make it easier to remember where each column came from.
```{r}
names(surveys_wide_symptomatic_outcome) <- paste0("outcome_",names(surveys_wide_symptomatic_outcome))
names(hrv_covid19_tables$participants.csv) <- paste0("participants_",names(hrv_covid19_tables$participants.csv) )
names(sleep_clean) <- paste0("sleep_",names(sleep_clean))
names(hrv_covid19_tables$wearables.csv) <- paste0("wearables_",names(hrv_covid19_tables$wearables.csv))
names(hrv_covid19_tables$weather.csv) <- paste0("weather_",names(hrv_covid19_tables$weather.csv))
names(hrv_measurements_clean_summarised) <- paste0("hrv_",names(hrv_measurements_clean_summarised))
names(blood_pressure_clean_summarised) <- paste0("bloodpressure_",names(blood_pressure_clean_summarised) )
names(heart_rate_clean_summarised) <- paste0("heartrate_",names(heart_rate_clean_summarised))
```



```{r}
surveys_wide_join <- left_join(surveys_wide_symptomatic_outcome,hrv_covid19_tables$participants.csv, by = c("outcome_user_code" = "participants_user_code"))

surveys_wide_join <- left_join(surveys_wide_join,sleep_clean, by = c("outcome_user_code" = "sleep_user_code",
                                                                                                            "outcome_created_at" = "sleep_day"))

surveys_wide_join <- left_join(surveys_wide_join,hrv_covid19_tables$wearables.csv, by = c("outcome_user_code" = "wearables_user_code",
                                                                                                            "outcome_created_at" = "wearables_day"))

surveys_wide_join <- left_join(surveys_wide_join,hrv_covid19_tables$weather.csv, by = c("outcome_user_code" = "weather_user_code",
                                                                                                            "outcome_created_at" = "weather_day"))

surveys_wide_join <- left_join(surveys_wide_join,hrv_measurements_clean_summarised, by = c("outcome_user_code" = "hrv_user_code",
                                                                                                            "outcome_created_at" = "hrv_day"))

surveys_wide_join <- left_join(surveys_wide_join,blood_pressure_clean_summarised, by = c("outcome_user_code" = "bloodpressure_user_code",
                                                                                                            "outcome_created_at" = "bloodpressure_day"))

surveys_wide_join <- left_join(surveys_wide_join,heart_rate_clean_summarised, by = c("outcome_user_code" = "heartrate_user_code",
                                                                                                            "outcome_created_at" = "heartrate_day"))

```
## Convert relevant variables as factors


Convert characters to factors

```{r}
surveys_wide_join <- surveys_wide_join %>%
  mutate_if(is.character, as.factor)
```

Convert ordinal outcome to ordered

```{r}
surveys_wide_join <- surveys_wide_join %>%
  mutate(outcome_S_COVID_OVERALL = as.ordered(outcome_S_COVID_OVERALL))

```

# Split
```{r}
set.seed(42)
train_idx <- caret::createDataPartition(y = surveys_wide_join$outcome_symptom_severity, p = .8, list = F)
surveys_wide_join_train_data <- surveys_wide_join[train_idx,]
surveys_wide_join_test_data <- surveys_wide_join[-train_idx,]
```

```{r eval = FALSE}
write_tsv(surveys_wide_join_train_data,"../data_clean/surveys_wide_join_train_data.tsv")
write_tsv(surveys_wide_join_test_data,"../data_clean/surveys_wide_join_test_data.tsv")

```

# Clean data  

# Pre-process

Remove variables not to use in the model 
```{r}
surveys_wide_join_clean <- surveys_wide_join %>%
  dplyr::select(- participants_symptoms_onset, - participants_country, - participants_city)
```

Re-code outcome to ordered
```{r}
surveys_wide_join_clean <- surveys_wide_join_clean %>%
  mutate(outcome_S_COVID_OVERALL = as.ordered(outcome_S_COVID_OVERALL))
```


Explore NA values
```{r}
na_proportion_per_feature <- (map_dbl(surveys_wide_join_clean, function(x) sum(is.na(x))) / nrow(surveys_wide_join_clean)) 


na_proportion_per_feature %>%
  enframe() %>%
  ggplot(aes(value)) +
  geom_histogram() +
  theme_light() +
  theme(aspect.ratio = 1) +
  ggtitle("Proportion of NA values per feature")
```

Remove features with > 60% NA values
```{r}
surveys_wide_join_clean <- surveys_wide_join_clean %>%
  select_if(na_proportion_per_feature < 0.6)
```

Just use mean imputation for now

Start with per person mean imputation

```{r}
surveys_wide_join_clean <- surveys_wide_join_clean %>%
  group_by(outcome_user_code) %>%
  mutate_if(is.numeric,~ ifelse(is.na(.),mean(., na.rm = TRUE),.)) 
```

If a person does not have a value, then impute by global mean
```{r}
surveys_wide_join_clean <- surveys_wide_join_clean %>%
  ungroup() %>%
  mutate_if(is.numeric,~ ifelse(is.nan(.),mean(., na.rm = TRUE),.)) 
```

If a person does not have a value, then impute by global mean
```{r}
surveys_wide_join_clean <- surveys_wide_join_clean %>%
  mutate_if(is.numeric,~ ifelse(is.infinite(.),mean(., na.rm = TRUE),.)) 

```

Many inf values in rr data, so drop for now
```{r}
sapply(surveys_wide_join_clean,function(x) any(is.infinite(x)))

```

```{r}
surveys_wide_join_clean <- surveys_wide_join_clean %>%
  dplyr::select(-hrv_rr_data)

```

Correlations between variables
```{r}
corrplot::corrplot(cor(surveys_wide_join_clean %>%
      select_if(is.numeric)))
```

Remove correlated vars 

```{r}
hrv_vars_to_keep <- surveys_wide_join_clean %>%
  dplyr::select(hrv_total_power,hrv_how_feel,hrv_how_mood, hrv_how_sleep, hrv_vlf, hrv_lfhf)

surveys_wide_join_clean <- surveys_wide_join_clean %>%
  dplyr::select(!starts_with("hrv"))
surveys_wide_join_clean <- cbind(surveys_wide_join_clean, hrv_vars_to_keep)

```

## Split train / test

```{r}
set.seed(42)
train_idx <- caret::createDataPartition(y = surveys_wide_join_clean$outcome_symptom_severity, p = .8, list = F)
surveys_wide_join_train_data_clean <- surveys_wide_join_clean[train_idx,]
surveys_wide_join_test_data_clean <- surveys_wide_join_clean[-train_idx,]
```

```{r eval = FALSE}
write_tsv(surveys_wide_join_train_data_clean,"../data_clean/surveys_wide_join_train_data_clean.tsv")
write_tsv(surveys_wide_join_test_data_clean,"../data_clean/surveys_wide_join_test_data_clean.tsv")

```


# Alternative data table, for unsupervised approach purposes

Joining full hrv data table w/o summarization to survey and other participant data.  

```{r}
hrv_measurements_clean_full <- hrv_covid19_tables$hrv_measurements.csv %>%
  separate(measurement_datetime, into = c("day","time"), sep= " ") %>%
  mutate(day = as.Date(day))

```

```{r}
colnames(surveys_wide) <- paste0("surveys_", colnames(surveys_wide))
colnames(hrv_measurements_clean_full) <- paste0("hrv_", colnames(hrv_measurements_clean_full))

```


Join surveys and participant data (summarized per user /day)
```{r}
surveys_hrv_join <- left_join(surveys_wide,hrv_covid19_tables$participants.csv, by = c("surveys_user_code" = "participants_user_code"))

surveys_hrv_join <- left_join(surveys_hrv_join,sleep_clean, by = c("surveys_user_code" = "sleep_user_code",
                                                                                                            "surveys_created_at" = "sleep_day"))

surveys_hrv_join <- left_join(surveys_hrv_join,hrv_covid19_tables$wearables.csv, by = c("surveys_user_code" = "wearables_user_code",
                                                                                                            "surveys_created_at" = "wearables_day"))

surveys_hrv_join <- left_join(surveys_hrv_join,hrv_covid19_tables$weather.csv, by = c("surveys_user_code" = "weather_user_code",
                                                                                                            "surveys_created_at" = "weather_day"))

```

Join 

```{r}

surveys_hrv_join <- left_join(hrv_measurements_clean_full,surveys_hrv_join, by = c("hrv_user_code" = "surveys_user_code",
                                                                                                "hrv_day" = "surveys_created_at"))
```


```{r}
write_tsv(surveys_hrv_join,"../data_clean/surveys_hrv_join.tsv")

```


```{r}
library(tidyverse)

surveys_hrv_join <- read_tsv("../data_clean/surveys_hrv_join.tsv")

head(surveys_hrv_join)
```

